Plots for Latent Hawkes Paper

### 2020-10-12 output_scripts/Paper_Plots.Rmd
start putting together a new plot for fig 3b

ks.stat.matrix <- matrix(0,ncol=mice_number,nrow=mice_number)
n_matrix <- matrix(0,ncol=mice_number,nrow=mice_number)
Lambda_matrix <- matrix(list(),nrow=mice_number,ncol=mice_number)
for(k in 1:mice_number){
  for(l in c(1:mice_number)[-k]){
    i<-expert_rank_10[[current_cohort]][k]
    j<-expert_rank_10[[current_cohort]][l]
    pair <- which(unique_pairs_df$initiator==i&unique_pairs_df$recipient==j)
    if(length(pair)>0){
      param_mean <- list(lambda0=mean(sim_cohort_hp$lambda0),
                         alpha=model1_fn$alpha.fun(f_est[i],f_est[j],
                                                   eta1_mean,eta2_mean,eta3_mean),
                         beta=mean(sim_cohort_hp$beta))
      current_window_vec <- unique_pairs_df$observe[[pair]]
      all_Lambda <- numeric(0)
      for(cur in c(1:length(current_window_vec))){ ## check length > 2
        cur_win <- current_window_vec[cur]
        current_event_time <- return_df[return_df$initiator==i&
                                          return_df$recipient==j&
                                          return_df$observe.id==cur_win,"event.times"][[1]]
        current_obs_time <- return_df[return_df$initiator==i&
                                        return_df$recipient==j&
                                        return_df$observe.id==cur_win,"observe.time"]
        Lambda_vec <- uniHawkesCompensator(object=param_mean, events=c(current_event_time,current_obs_time))
        all_Lambda <- c(all_Lambda,Lambda_vec)
        n_matrix[i,j] <- n_matrix[i,j] + length(current_event_time)
      }
      if(length(all_Lambda)>0){
        ks.stat.matrix[i,j] <-  ks.test(all_Lambda,"pexp",rate=1/mean(all_Lambda))$statistic
        Lambda_matrix[i,j][[1]] <- all_Lambda
      }
    }
  }
}

@@

####
par(oma = c(6,4,4,1), mar=rep(0,4),
    mgp = c(0, .2, 0))
layout(cbind(matrix(data=c(1:(12*12)), nrow=12, ncol=12, byrow = TRUE),rep(12*12+1,12)), 
       widths=c(rep(1,12),2), heights=rep(1,12))
for(i in expert_rank_10[[current_cohort]]){
  for(j in expert_rank_10[[current_cohort]]){
    if(ks.stat.matrix[i,j]==0){
      plot(0,0,xlim=c(0,5),ylim=c(0,5),type="n", xlab="", ylab="", axes = FALSE, frame=TRUE)
      rect(par('usr')[1],par('usr')[3],par('usr')[2],par('usr')[4], border = 0, col = 'white')
    }else{
      Lambda.test <- Lambda_matrix[i,j][[1]]
      curve(pexp(x,1/mean(Lambda.test)), min(Lambda.test), max(Lambda.test),
            ylab="",xlab="",axes = FALSE, frame=TRUE)
      rect(par('usr')[1],par('usr')[3],par('usr')[2],par('usr')[4], border = 0, col = col.vec.vir[matrixToVecIdx(i,j,mice_number)])
      curve(pexp(x,1/mean(Lambda.test)), min(Lambda.test), max(Lambda.test),
            ylab="",xlab="",axes = FALSE, frame=TRUE,add=TRUE, lwd=2)
      plot(ecdf(Lambda.test), add=TRUE, do.points=FALSE, verticals=TRUE,col="red", lwd=2)
    }
  }
}
par(mar = c(0,2,0,3),mgp=c(0,0.8,0))

@@

mtext(text=expert_rank_10[[current_cohort]], side=1, line=1.8, outer=TRUE, cex=1.8,
      at=1/14*c(1:12)-1/22)
mtext(text=expert_rank_10[[current_cohort]], side=2, line=0.5, outer=TRUE, cex=1.8,
      at=1/12*c(12:1)-1/20, las=2)
mtext(text="(b)", side=3, line=1.4, outer=TRUE, cex=2.2,
      at=0.45, font=2)
#dev.off()
```

@@

# social matrix plot
latent_social_matrix <- gamma_mean%*%t(rep(1,mice_number))+rep(1,mice_number)%*%t(zeta_mean)

# png(paste(plot_path,"/lambda0_",cohort_names[current_cohort],".png",sep=''),
#     width=400, height=300)
matrixPlotParameter(latent_social_matrix[rev(expert_rank_10[[current_cohort]]),
                                         expert_rank_10[[current_cohort]]],
                    xLabels=expert_rank_10[[current_cohort]],
                    yLabels=rev(expert_rank_10[[current_cohort]]),title="(a)")
# mtext(text=c(1,2,11,6,7,9), side=1, line=-0.35, outer=TRUE, cex=1.3,
#       at=1/17.8*seq(2,12,2)+0.09)
#dev.off()
```
## Figure 3b: Ranking Comparison

@@

@@

```{r fig-10b-rank-correlation}
par(mar=c(3,5,3,1))
cx <- 2;  # cex for axis label text; default is 1. bigger numbers make the text bigger, smaller, smaller.  
x.ttl <- ""; # blank x-axis label  
y.ttl <- "rank correlation (expert rank)";  
ttl <- "(b)";  
#y.lim <- c(min(data.tbl$data.value), max(data.tbl$data.value)+.5); # y-axis limits. +2 at top to give room for legend  
y.lim <- c(-1,1.2)    
x.lim <- c(0.5, (length(task.types))) 
cols <- c('paleturquoise4', 'lightgreen', 'cornsilk')
shifts <- c(-0.25, 0, 0.25)

plot(x=0, y=0, xlim=x.lim, ylim=y.lim, xaxt='n', col='white', xlab=x.ttl, ylab=y.ttl, main=ttl, cex.main=cx*1.3, cex.axis=cx, cex.lab=cx); 
axis(side=1, at=1:length(task.types), labels=task.types, cex.axis=cx, cex.lab=cx)  # put on the x-axis labels  
grid(nx=NA, ny=NULL, col='darkgrey');  
lines(x=c(-1,100), y=c(0,0), col='darkgrey');  
for (t.id in 1:length(task.types)) {  
  for (i in 1:length(sub.types)) {  # t.id <- 1; i <- 1;   
   inds <- which(data.tbl$sub.type == sub.types[i] & data.tbl$task.type == task.types[t.id])
   boxplot(data.tbl$data.value[inds], at=t.id+shifts[i], col=cols[i], add=TRUE, 
           xaxt='n', yaxt='n', bty='n', boxwex=0.35, outpch = 20, outcex = 0.5) 
  }  
}  
legend(x='top', legend=sub.types, fill=cols, horiz=TRUE, cex=cx, bg='white', bty='n')  
box()
```

>>>>

ks.stat.matrix <- matrix(0,ncol=mice_number,nrow=mice_number)
n_matrix <- matrix(0,ncol=mice_number,nrow=mice_number)
Lambda_matrix <- matrix(list(),nrow=mice_number,ncol=mice_number)
for(k in 1:mice_number){
  for(l in c(1:mice_number)[-k]){
    i<-expert_rank_10[[current_cohort]][k]
    j<-expert_rank_10[[current_cohort]][l]
    pair <- which(unique_pairs_df$initiator==i&unique_pairs_df$recipient==j)
    if(length(pair)>0){
      param_mean <- list(lambda0=mean(sim_cohort_hp$lambda0),
                         alpha=model1_fn$alpha.fun(f_est[i],
                                                   f_est[j],
                                                   eta1_mean,
                                                   eta2_mean,
                                                   eta3_mean),
                         beta=mean(sim_cohort_hp$beta))
      current_window_vec <- unique_pairs_df$observe[[pair]]
      all_Lambda <- numeric(0)
      for(cur in c(1:length(current_window_vec))){ ## check length > 2
        cur_win <- current_window_vec[cur]
        current_event_time <- return_df[return_df$initiator==i&
                                          return_df$recipient == j &
                            return_df$observe.id==cur_win,"event.times"][[1]]
        current_obs_time <- return_df[return_df$initiator==i&
                                        return_df$recipient==j&
                            return_df$observe.id==cur_win,"observe.time"]
        Lambda_vec <- uniHawkesCompensator(object = param_mean,
                              events = c(current_event_time,
                                         current_obs_time))
        all_Lambda <- c(all_Lambda,Lambda_vec)
        n_matrix[i,j] <- n_matrix[i,j] + length(current_event_time)
      }
      if(length(all_Lambda)>0){
        ks.stat.matrix[i,j] <-  ks.test(all_Lambda,"pexp",
                                        rate = 1/mean(all_Lambda))$statistic
        Lambda_matrix[i,j][[1]] <- all_Lambda
      }
    }
  }
}

@@

####
par(oma = c(6,4,4,1), mar=rep(0,4),
    mgp = c(0, .2, 0))
layout(cbind(matrix(data=c(1:(12*12)), nrow = 12, ncol = 12, 
                    byrow = TRUE),rep(12*12+1,12)), 
       widths=c(rep(1,12),2), heights=rep(1,12))
for(i in expert_rank_10[[current_cohort]]){
  for(j in expert_rank_10[[current_cohort]]){
    if(ks.stat.matrix[i,j]==0){
      plot(0,0,xlim=c(0,5),ylim=c(0,5),type="n", xlab="", ylab="", 
           axes = FALSE, frame=TRUE)
      rect(par('usr')[1],par('usr')[3],par('usr')[2],par('usr')[4],
           border = 0, col = 'white')
    }else{
      Lambda.test <- Lambda_matrix[i,j][[1]]
      curve(pexp(x,1/mean(Lambda.test)), min(Lambda.test), max(Lambda.test),
            ylab="",xlab="",axes = FALSE, frame=TRUE)
      rect(par('usr')[1],par('usr')[3],par('usr')[2],par('usr')[4], 
           border = 0, col = col.vec.vir[matrixToVecIdx(i,j,mice_number)])
      curve(pexp(x,1/mean(Lambda.test)), min(Lambda.test), max(Lambda.test),
            ylab="",xlab="",axes = FALSE, frame=TRUE,add=TRUE, lwd=2)
      plot(ecdf(Lambda.test), add = TRUE, do.points = FALSE,
           verticals = TRUE, col = "red", lwd = 2)
    }
  }
}
par(mar = c(0,2,0,3),mgp=c(0,0.8,0))

@@

mtext(text=expert_rank_10[[current_cohort]], side=1, line=1.8, outer=TRUE, cex=1.8,
      at=1/14*c(1:12)-1/22)
mtext(text = expert_rank_10[[current_cohort]],
      side = 2, line = 0.5, outer = TRUE, cex = 1.8,
      at =1/12*c(12:1)-1/20, las = 2)
mtext(text = "(b)", side = 3, line = 1.4, outer = TRUE,
      cex = 2.2,
      at=0.45, font=2)
#dev.off()
```

@@

# social matrix plot
latent_social_matrix <- gamma_mean %*%
  t(rep(1,mice_number)) + rep(1,mice_number) %*% t(zeta_mean)

# png(paste(plot_path,"/lambda0_",
# cohort_names[current_cohort],".png",sep=''),
#     width=400, height=300)
matrixPlotParameter(latent_social_matrix[
  rev(expert_rank_10[[current_cohort]]),
                                         expert_rank_10[[current_cohort]]],
                    xLabels=expert_rank_10[[current_cohort]],
                    yLabels=rev(expert_rank_10[[current_cohort]]),
                    title = "(a)" )
# mtext(text=c(1,2,11,6,7,9), side=1, line=-0.35, outer=TRUE, cex=1.3,
#       at=1/17.8*seq(2,12,2)+0.09)
#dev.off()
```
## Figure 3b: Ranking Comparison

@@

# Redone Figure 3

As discussed, it would be interesting to redo figure 3b
to show the lack of fit in model 2 for one specific pair. This can be somewhat
extracted from the diagnostics file.

We actually want the empirical CDF plot here.

```{r fig_3b_update, fig.height=7}
# 2 -> 11 might be good here
current_cohort <- 5
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events)) 
top_pairs <- unique_pairs_df %>% arrange(desc(count))

best_i <- top_pairs$initiator[8]
best_j <- top_pairs$recipient[8]

current_i <- best_i #8
current_j <- best_j #2

### then plots ####
# par(mfrow = c(1, 2), oma = c(1, 1, 1, 1),
#      mar = c(2, 3, 2, 1), mgp = c(0.8,0,0))
# layout(matrix(1:8,nrow=2,byrow=TRUE),widths=rep(2,4),height=c(3,3))
anno.cex <- 1.3
title.cex <- 2.2
title.line <- 0.5
y.upper <- 20
#### then the corresponding QQ plot ####
load(paste(data_path,cohort_names[current_cohort],
             "/real_diag_four_models_",cohort_names[current_cohort],
             ".RData",sep=''))
Lambda.dc_hawkes.test <- Lambda_dc_hawkes_matrix[current_i,current_j][[1]]

p <- ppoints(length(Lambda.dc_hawkes.test))
# 100 equally spaced points on (0,1), excluding endpoints
q <- quantile(Lambda.dc_hawkes.test,p=p,na.rm=TRUE)
# percentiles of the sample distribution
plot(qexp(p),q,
     xlab="",ylab="",xaxt="n",yaxt="n",
     pch=1, cex=1, col=viridis::viridis(1), ylim=c(0,1), xlim=c(0,4.5))
mtext(side = 1, text = "Theoretical", line=1.6, cex=anno.cex)
mtext(side = 2, text = "Empirical", line = 1.2, cex = anno.cex)
mtext(side = 3, text = "(b)", line = 0.8, cex = 1.6, font = 2)
# Hmisc::mgp.axis(1,mgp=c(0,0.1,0),at=c(0:5),cex.axis=1,lwd=0,lwd.ticks=1)
# Hmisc::mgp.axis(2,mgp=c(0,0.1,0),at=c(0:1),cex.axis=1,lwd=0,lwd.ticks=1)
qqline(q, distribution=qexp,col="Black", lty=1, lwd=2.5)

### Empirical CDF ####

plot(ecdf(Lambda.dc_hawkes.test), do.points = FALSE,
     verticals = TRUE, col = viridis::viridis(1), lwd = 2, 
     xlab = "", ylab = "", xaxt = "n", yaxt = "n",
     main = "", col.01line = NULL)
curve(pexp(x,1/mean(Lambda.dc_hawkes.test)),
      min(Lambda.dc_hawkes.test), max(Lambda.dc_hawkes.test),
      ylab = "", xlab = "", axes = FALSE,
      add = TRUE, lwd = 2)

```{r repeat entire fig2b for model 2}
current_cohort <- 5

load(paste(data_path,cohort_names[current_cohort],
             "/real_diag_four_models_",cohort_names[current_cohort],
             ".RData",sep=''))

load(paste(data_path,cohort_names[current_cohort],
           "/cohort_dchp_stan_result_",cohort_names[current_cohort],
           ".RData",sep=''))
eta1_mean <- mean(sim_cohort_dchp$eta_1)
eta2_mean <- mean(sim_cohort_dchp$eta_2)
eta3_mean <- mean(sim_cohort_dchp$eta_3)
f_est <- apply(sim_cohort_dchp$f,2,mean)
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
                   observe=list(observe.id),
                   observe.length=list(observe.time),
                   no.events=list(no.events))
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

ks.stat.matrix <- matrix(0,ncol=mice_number,nrow=mice_number)
n_matrix <- matrix(0,ncol=mice_number,nrow=mice_number)
Lambda_matrix <- matrix(list(),nrow=mice_number,ncol=mice_number)
for(k in 1:mice_number){
  for(l in c(1:mice_number)[-k]){
    i<-expert_rank_10[[current_cohort]][k]
    j<-expert_rank_10[[current_cohort]][l]
    pair <- which(unique_pairs_df$initiator==i&unique_pairs_df$recipient==j)
    if(length(pair)>0){
      param_mean <- list(lambda0=mean(sim_cohort_dchp$lambda0[,pair]),
                         alpha=model1_fn$alpha.fun(f_est[i],
                                                   f_est[j],
                                                   eta1_mean,
                                                   eta2_mean,
                                                   eta3_mean),
                         beta=mean(sim_cohort_dchp$beta))
      current_window_vec <- unique_pairs_df$observe[[pair]]
      all_Lambda <- numeric(0)
      for(cur in c(1:length(current_window_vec))){ ## check length > 2
        cur_win <- current_window_vec[cur]
        current_event_time <- return_df[return_df$initiator==i&
                                          return_df$recipient == j &
                            return_df$observe.id==cur_win,"event.times"][[1]]
        current_obs_time <- return_df[return_df$initiator==i&
                                        return_df$recipient==j&
                            return_df$observe.id==cur_win,"observe.time"]
        Lambda_vec <- uniHawkesCompensator(object = param_mean,
                              events = c(current_event_time,
                                         current_obs_time))
        all_Lambda <- c(all_Lambda,Lambda_vec)
        n_matrix[i,j] <- n_matrix[i,j] + length(current_event_time)
      }
      if(length(all_Lambda)>0){
        ks.stat.matrix[i,j] <-  ks.test(all_Lambda,"pexp",
                                        rate = 1/mean(all_Lambda))$statistic
        Lambda_matrix[i,j][[1]] <- all_Lambda
      }
    }
  }
}

# -------------- KS plot (stat)
col.vec <- cRamp(as.vector(ks.stat.matrix))
## need to do some sort of transformation between these here
vir_cols <- viridis::viridis(length((as.vector(ks.stat.matrix))))
#order(as.vector(ks.stat.matrix),method = "")

####
par(oma = c(6,4,4,1), mar=rep(0,4),
    mgp = c(0, .2, 0))
layout(cbind(matrix(data=c(1:(12*12)), nrow = 12, ncol = 12, 
                    byrow = TRUE),rep(12*12+1,12)), 
       widths=c(rep(1,12),2), heights=rep(1,12))
for(i in expert_rank_10[[current_cohort]]){
  for(j in expert_rank_10[[current_cohort]]){
    if(ks.stat.matrix[i,j]==0){
      plot(0,0,xlim=c(0,5),ylim=c(0,5),type="n", xlab="", ylab="", 
           axes = FALSE, frame=TRUE)
      rect(par('usr')[1],par('usr')[3],par('usr')[2],par('usr')[4],
           border = 0, col = 'white')
    }else{
      Lambda.test <- Lambda_matrix[i,j][[1]]
      curve(pexp(x,1/mean(Lambda.test)), min(Lambda.test), max(Lambda.test),
            ylab="",xlab="",axes = FALSE, frame=TRUE)
      rect(par('usr')[1],par('usr')[3],par('usr')[2],par('usr')[4], 
           border = 0, col = col.vec.vir[matrixToVecIdx(i,j,mice_number)])
      curve(pexp(x,1/mean(Lambda.test)), min(Lambda.test), max(Lambda.test),
            ylab="",xlab="",axes = FALSE, frame=TRUE,add=TRUE, lwd=2)
      plot(ecdf(Lambda.test), add = TRUE, do.points = FALSE,
           verticals = TRUE, col = "red", lwd = 2)
    }
  }
}
par(mar = c(0,2,0,3),mgp=c(0,0.8,0))

col_obj <- list(x.vec = seq(0,max(as.vector(ks.stat.matrix)),
                            length.out = 100),
                col.vec = viridis(100))

temp <- cLegend(as.vector(ks.stat.matrix))

image(1, col_obj$x.vec,
      matrix(data=col_obj$x.vec, 
             ncol=length(col_obj$x.vec),nrow=1),
      col=col_obj$col.vec,
      xlab="",ylab="",axes=FALSE)
axis(4, at=c(0,0.3,0.6), labels=c(0,0.3,0.6), las= HORIZONTAL<-1,
     cex.axis=2.4)

mtext(text=expert_rank_10[[current_cohort]], side=1,
      line=1.8, outer=TRUE, cex=1.8,
      at=1/14*c(1:12)-1/22)
mtext(text = expert_rank_10[[current_cohort]],
      side = 2, line = 0.5, outer = TRUE, cex = 1.8,
      at =1/12*c(12:1)-1/20, las = 2)
mtext(text = "(b)", side = 3, line = 1.4, outer = TRUE,
      cex = 2.2,
      at=0.45, font=2)

This is good because it shows the most active animal, shows that individual
out rate too large for this interaction. Just need to make this more
like the plots in 2b.

@@

```{r fig-10b-rank-correlation}
par(mar=c(3,5,3,1))
cx <- 1.75  # cex for axis label text; 
# default is 1. bigger numbers make the text bigger, smaller, smaller.  
x.ttl <- ""; # blank x-axis label  
y.ttl <- "rank corr. (expert rank)";  
ttl <- "(b)";  
#y.lim <- c(min(data.tbl$data.value), max(data.tbl$data.value)+.5);
# y-axis limits. +2 at top to give room for legend  
y.lim <- c(-1,1.2)    
x.lim <- c(0.5, (length(task.types))) 
cols <- c('paleturquoise4', 'lightgreen', 'cornsilk')
shifts <- c(-0.25, 0, 0.25)

plot(x = 0, y = 0, xlim = x.lim, ylim = y.lim, xaxt='n',
     col = 'white', xlab = x.ttl,
     ylab = y.ttl, main = ttl,
     cex.main = cx*1.3, cex.axis = cx, cex.lab = cx); 
axis(side = 1, at = 1:length(task.types), labels = task.types,
     cex.axis=cx, cex.lab=cx)  # put on the x-axis labels  
grid(nx=NA, ny=NULL, col='darkgrey');  
lines(x=c(-1,100), y=c(0,0), col='darkgrey');  
for (t.id in 1:length(task.types)) {  
  for (i in 1:length(sub.types)) {  # t.id <- 1; i <- 1;   
   inds <- which(data.tbl$sub.type == sub.types[i] & 
                   data.tbl$task.type == task.types[t.id])
   boxplot(data.tbl$data.value[inds], at=t.id+shifts[i],
           col=cols[i], add=TRUE, 
           xaxt='n', yaxt='n', bty='n', boxwex=0.35,
           outpch = 20, outcex = 0.5) 
  }  
}  
legend(x='top', legend = sub.types, fill=cols,
       horiz = TRUE, cex = cx, bg = 'white', bty = 'n')  
box()
```
